---
title             : "Does varying cue-stimulus interval affect the sensory discrimination performance in the antisaccade task?"
shorttitle        : "Attention variance over time"

author: 
  - name          : "Bartłomiej Kroczek"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    # address       : "Postal address"
    email         : "bartek.kroczek@doctoral.uj.edu.pl"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Formal Analysis"
      - "Investigation"
      - "Methodology"
      - "Software"
      - "Visualization"
      - "Validation"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Adam Chuderski"
    affiliation   : "2"
    role:
      - "Conceptualization"
      - "Methodology"
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Institute of Psychology, Jagiellonian University, Krakow, Poland"
  - id            : "2"
    institution   : "Department of Cognitive Science, Jagiellonian University, Krakow, Poland"

authornote: |

  This manuscript is the doctoral thesis of Bartłomiej Kroczek, prepared under the supervision of professor Adam Chuderski, implemented as part of the CogNes 19 doctoral program.
  
  All code and data used in performing this research are publicly available at `https://github.com/bartekkroczek/PhD`

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "attentional vigilance, antisaccade task, processing speed"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : yes
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Rozdział 1 
Tu mam najtrudniej, bo złapałem się na luce w rozumieniu. Jaki - z naszej perspektywy - jest związek między attentional vigilance a cognitive control? Czemu torturując zadanie antysakadowe które ma być miarą kontroli wysnuwamy wnioski o uwadze? To jest tak, że cognitive control to jakiś puzel, który służy do budowania teorii jak działa uwaga (i pewnie nie tylko)? 

Pomijając powyższe, wyobrażam sobie takie fakty:

* Argumenty za tym, że uwaga jest zmienna w czasie (bo mamy np. mrugnięcie uwagowe, saccadic supression, oscylacyjne teorie wszystkiego itp).
* Wykazanie, że zadanie antysakadowe jest powszechnie poważane jako dobra miara kontroli 
* Brakujący mi klej między kontrolą a uwagą
* Argumenty literaturowe za istnieniem trendu i oscylacji (behawioralnych) w varying cue-stimuli interval
* Pointa, że to się klei, że skoro uwaga ma być niestabilna w czasie, to i zadanie je mierzące powinno być niestabilne w czasie i to właśnie pokazuje literatura. (Znów, brak kleju)

# Rozdział 2
Zarysowanie problemu, a mianowicie, że chcemy na własne oczy sprawdzić co to się dzieje w tym varying cue-stimuli interval. 
Tutaj nie wiem na ile warto/należy spoilerować zakończenie.

# Chapter III. What our data shows? 

To examine the role of CSI in stimulus discrimination, we used the antisaccade 
task with CSI varied in millisecond steps. On each trial, participants were shown for 250 ms
either left or right arrow as a target, and their task was to press the appropriate
key. The arrows were randomly displayed on either the left or right side of the
screen. The stimulus was accompanied by a red dot briefly flashing on the
opposite side of the screen, which should be ignored. The fixation point and
the red dot/stimulus presentation were separated by a blank screen shown for
CSI ranging from 400 to 900ms. We conducted two experiments, one with lower
resolution (less trials per participant) but a larger sample (N=150, CSI sampling
frequency 60Hz) and the other with higher resolution but a smaller sample
(N=40, CSI sampling frequency 120Hz). Both data sets were analyzed in the
same way. For each participant, the mean accuracy was calculated for each CSI
timepoint. These mean accuracy values were plotted as a function of increasing
CSI, with a single curve created for each person.


## Anti-saccade task

The antisaccade task is one of the most popular measures of cognitive control.
On each trial, participants were shown for 250 ms either left, right, 
up or down arrow as a target, and their task was to press the appropriate key.
The arrows were randomly displayed on either the left or right side of the screen.
The stimulus was accompanied by a red dot briefly flashing on the opposite side
of the screen, which should be ignored. The fixation point at the very beginning 
of a trial, and the red dot/stimulus presentation were separated by a blank screen
shown for a ranging time, called the cross-stimuli interval (CSI).  

## Data preparation


## Sekcja 3.1 
Dane N=150, niskie próbkowanie CSI 200-1000 ms. Już zebrane i przeanalizowane. 
  Wniosek - jest trend, ale malutki, dużo mniejszy niż w literaturze (i hope so)
    
### Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Sekcja 3.2
Dane N=40 wysokie próbkowanie, CSI 400-900 ms. Już zebrane i przeanalizowane. Wniosek, brak trendu, zarówno w analizie zbiorczej dla każdej z częstotliwości (to od Tomasza) jak i różnic indywidualnych (kończę robić)
    
## Sekcja 3.3 
Dane N=Przyzwoicie_Ale_Nie_Za_Duzo niskie próbkowanie, CSI 500(?)-3000(?)ms Nie zebrane. 
  Wniosek, daj Bóg, że trend znika.

# Rozdział 4 

Dlaczego nasze wyniki nie pokrywają się z literaturą? 
    Sekcja 4.1 Trend
    Bo efekt wisiał na punkcie w 200 ms czyli na mrugnięciu uwagowym.
    Sekcja 4.2 Oscylacje
    Wnioski i symulacje z  artykułu metodologicznego o niepoprawnym preprocessigu. W końcu jest to już, marnie bo marnie, ale napisane. 
 
Konkluzja: Nasze dane pokazują, że cue-stimuli interval nie oscyluje i ma bardzo mały/nie ma efektu trendu. 

# Rozdział 5
Nasze obserwacje nie zgadzają się z obecnym stanem rzeczy. Co gorsza, twierdzimy że wiemy dlaczego. Nasi biografowie nazwą to zjawisko Kroczek&Chuderski Tension w analogii do Hubble Tension - największego kryzysu współczesnej kosmologii. W tym momencie czytelnik ma już mokre poty i zjadł wszystkie paznokcie. Gdy myśli, że jego nerwy nie mogą być już bardziej napięte to walimy go w łeb obuchem stawiając pytanie, czy w takim razie uwaga/kontrola jest jednak stała w czasie, czy zadanie antysakadowe nie jest dobrą miarą kontroli. Czytelnik mdleje.

# Rozdział 6 

W oparciu o Roudera stawiamy tezę, że zadanie antysakadowe nie mierzy kontroli i przedstawiamy kolejny eksperyment który temu dowodzi. 
Opisujemy eksperyment i pokazujemy, że zadanie antysakadowe nie tyle jest marną miarą kontroli, co jest zadaniem na processing speed. 

# Rozdział 7 
Konkluzja pracy, w jakich obszarach (teorie uwagi) nasze wnioski mogą namieszać. Future directions. Tłum wiwatuje w ekstazie. Kurtyna. 


<!-- We used `r cite_r("r-references.bib")` for all our analyses. -->



<!-- \newpage -->

# References

::: {#refs custom-style="Bibliography"}
:::
